---
title: "p8105_hw6_gp2770"
author: "Gokul Pareek"
date: "2024-11-19"
output: github_document
---

I'm an R Markdown document!

```{r}

library(tidyverse)
library(broom)

```

# Problem 1

## Data Generation

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

## Bootstrap Analysis

```{r}

set.seed(123)

# Function to compute R^2 and log(beta0 * beta1)
compute_metrics <- function(data) {
  model <- lm(tmax ~ tmin, data = data)
  r_squared <- glance(model)$r.squared
  coefs <- tidy(model)
  log_beta_product <- log(coefs$estimate[1] * coefs$estimate[2])
  c(r_squared, log_beta_product)
}

# Perform bootstrap
n_boot <- 5000
bootstrap_results <- replicate(n_boot, {
  sample_data <- weather_df %>% slice_sample(n = nrow(weather_df), replace = TRUE)
  compute_metrics(sample_data)
}, simplify = "matrix")

# Convert to a tidy data frame
bootstrap_df <- as.data.frame(t(bootstrap_results))
colnames(bootstrap_df) <- c("r_squared", "log_beta_product")

```

# Problem 2

```{r}

library(purrr)

```

## Loading and Modifying the Data

```{r}

homicide_df <- read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  # Clean column names for consistency
  janitor::clean_names() %>% 
  
  # Combine city and state into one column
  unite(city_state, c(city, state), sep = ", ") %>%
  
  # Filter out specific cities and keep only relevant races and valid ages
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black"),
    victim_age != "Unknown"
  ) %>%
  
  # Convert victim_age to numeric and create a resolved column as a binary indicator
  mutate(
    victim_age = as.numeric(victim_age),
    resolved = as.numeric(disposition == "Closed by arrest"),
    # Reorder levels in victim_race, prioritizing "White"
    victim_race = fct_relevel(victim_race, "White")
  ) %>%
  
  # Select only relevant columns for analysis
  select(city_state, resolved, victim_age, victim_race, victim_sex)

```

## Logistic Regression for Baltimore, MD

```{r}

logistic_model <- 
  homicide_df %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

broom::tidy(logistic_model, conf.int = TRUE) %>% mutate(OR = exp(estimate), 
                                                  conf_low = exp(conf.low), 
                                                  conf_high = exp(conf.high)) %>% 
  select(term, log_OR = estimate, OR, conf_low, conf_high) %>% filter(term == "victim_sexMale") %>% 
  knitr::kable(digits = 4)

```

## Logistic Regression for Each of the Cities

```{r}

city_results <- homicide_df %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = "binomial")),
    results = map(model, broom::tidy, conf.int = TRUE)) %>%
  unnest(results) %>%
  mutate(OR = exp(estimate), 
         conf_low = exp(conf.low), conf_high = exp(conf.high)) %>% 
  select(city_state, term, OR, conf_low, conf_high) %>%
   filter(term == "victim_sexMale") 

city_results %>% knitr::kable(digits = 4)

```

## Plotting

```{r}

city_results %>% 
  ggplot(aes(y = reorder(city_state, OR), x = OR)) +  # Cities on y-axis, OR on x-axis
  geom_point() +
  geom_errorbar(aes(xmin = conf_low, xmax = conf_high)) +  # Error bars for x-axis
  labs(
    title = "Unsolved Homicides: Male vs. Female Victims",
    y = "City",  # Reflects swapped axes
    x = "Adjusted Odds Ratio"
  ) +
  theme(
    axis.text.y = element_text(vjust = 0.5, margin = margin(t = 5, b = 5)),  # Adds spacing around labels
    axis.ticks.y = element_blank(),  # Optional: hides y-axis ticks for a cleaner look
    axis.text.x = element_text(angle = 0),  # Keeps x-axis text readable
    plot.title = element_text(hjust = 0.5)  # Centers the title
  )

```

# Problem 3

## Loading and Cleaning the Data

```{r}

library(modelr)

# Load the dataset
birthweight_data <- read.csv("data/birthweight.csv")

# Correct column name typo
colnames(birthweight_data)[colnames(birthweight_data) == "mheigth"] <- "mheight"

# Convert numeric columns to factors where appropriate
birthweight_data <- birthweight_data %>%
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4), 
                   labels = c("White", "Black", "Asian", "Puerto Rican"))
  )

# Check for missing data
missing_data_summary <- birthweight_data %>%
  summarize(across(everything(), ~ sum(is.na(.))))

# Display structure of the dataset to confirm changes
glimpse(birthweight_data)

```

## Proposing a Regression Model for Birthweight

Using hypothesized key predictors for birthweight:
1. Mother's weight (delivery weight, pre-pregnancy BMI)
2. Gestational age
3. Smoking during pregnancy
4. Baby's sex
5. Presence of malformations

```{r}

# Fit a hypothesized model
model_hypothesis <- lm(bwt ~ delwt + gaweeks + smoken + babysex + malform, data = birthweight_data)

# Add predictions and residuals to the dataset
birthweight_data <- birthweight_data %>%
  add_predictions(model_hypothesis, var = "predicted_bwt") %>%
  add_residuals(model_hypothesis, var = "residuals")

# Plot residuals vs. fitted values
ggplot(birthweight_data, aes(x = predicted_bwt, y = residuals)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals")

```

## Comparing Models

### Model 1: Length At Birth and Gestational Age 

```{r}

model_length_ga <- lm(bwt ~ blength + gaweeks, data = birthweight_data)

```

### Model 2: Full Interaction Model

```{r}

model_interactions <- lm(bwt ~ bhead * blength * babysex, data = birthweight_data)

```

### Cross-Validation for Comparison

```{r}

library(purrr)

cv_df <- 
  crossv_mc(birthweight_data, 100) %>% 
  mutate(
    train = map(train, as_tibble), 
    test = map(test, as_tibble)
  ) %>% 
  mutate(model_hypothesis = map(train, \(df) lm(bwt~ babysex + bhead + blength + delwt + gaweeks + mrace + parity + smoken, data = df)),
    model_length_ga = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)), 
         model_interactions = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex, data = df))) %>% 
  mutate(
    rmse_model_hypothesis = map2_dbl(model_hypothesis, test, \(mod, df) rmse(model = mod, data = df)), 
    rmse_model_length_ga = map2_dbl(model_length_ga, test, \(mod, df) rmse(model = mod, data = df)), 
    rmse_interactions = map2_dbl(model_interactions, test, \(mod, df) rmse(model = mod, data = df))
  )

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(everything(), 
               names_to = "model", 
               values_to = "rmse", 
               names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, fill = model)) + 
  geom_violin(trim = FALSE, color = "black") +
  scale_fill_manual(values = c("skyblue", "lightgreen", "salmon")) + # Custom colors
  theme_minimal() +
  labs(title = "RMSE for Different Models",
       x = "Model",
       y = "RMSE") +
  theme(legend.position = "none")

```







